---
title: "SS0206-15GC"
author: "Richard J. Telford"
date: "October 2, 2016"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("dplyr")
library("ggplot2")
```

```{r data}
dates <- read.table(header = TRUE, text = 
"depth_corr	Lab_Code	raw_14C_age	error
51	Poz-55131	4310	60
57	Poz-60276	4360	35
69.5	ETH56189	5268	102
71.5	Poz-51071	5420	50
74.5	ETH56190	5417	114
89.5	ETH56191	6495	147
93	Poz-60277	6450	50
95.5	ETH56962	6490	56
99	Poz-55133	6960	60
100.5	ETH56963	7165	107
102.5	Poz-57692	7650	40
108.5	Poz-55134	7830	50
118.5	Poz-55119	8240	60
129	Poz-60278	8690	50
135.4	ETH56964	9139	94
140.5	Poz-55120	9620	50
150.5	Poz-55121	10160	70
154.5	ETH56192	10292	241
161	Poz-59472	10630	60
170.5	Poz-57336	11380	60
180.5	Poz-59473	11900	110
206	Poz-55135	12690	100
216	Poz-59475	13130	130
226	Poz-55137	14190	100
240.5	ETH56965	13930	109
251.5	Poz-47990	14380	100
270.5	ETH56966	15093	152
290.5	Poz-57339	14950	70
300.5	Poz-59476	15480	90
312.5	Poz-51070	16010	100
352	Poz-56508	16670	70
406.5	Poz-47988	17730	130
440.5	Poz-55140	18690	120
490.5	Poz-55141	20150	220
535.5	Poz-51066	20930	170
")
```

Write OxCal model file
```{r}
  #calibrate singlely
ages <- paste0("R_date(\"Dep", dates$depth_corr, "\",", dates$raw_14C_age, ",", dates$error, ");")
ages <- paste(ages, collapse = "\n")

curve <- "Curve(\"marine13\");"
DR <- "Delta_R(\"Local Marine\",200,30);"

command <- paste(curve, DR, ages, sep = "\n")

writeLines(command, con = "oxcal.input")
system(paste("/home/gbsrt/oxcal/OxCal/bin/OxCalLinux", "oxcal.input"))
```

Import Oxcal function

```{r functions}
get.vectors <- function(js, target) {
  x <- js[grep(target, js)]
  lab <- gsub("^(ocd\\[\\d+\\]).*", "\\1", x )
  val <- gsub(".+=\\[(.+)\\];", "\\1", x)
  val <- strsplit(val, ", ")
  res <- sapply(val, as.numeric)
  setNames(res, lab)
}

get.values <- function(js, target, name = "value", numeric = TRUE) {
  x <- js[grep(target, js)]
  val <- gsub(".+=(.+);", "\\1", x)
  lab <- gsub("^(ocd\\[\\d+\\]).*", "\\1", x )
  if(numeric) {
    val <- as.numeric(val)
  } 
  res <- data_frame(lab, val)
  setNames(res, c("lab", name))
}
```

Import single dates
```{r}
js <- readLines("oxcal.js")
likelihoods <- get.vectors(js, "likelihood\\.prob=")
start <- get.values(js, "likelihood\\.start")
resolution <- get.values(js, "likelihood\\.resolution")

oxcal <- mapply(
  function(depth, prob, start, resolution, group){
    data.frame(depth = depth, date = start + 0:(length(prob) - 1) * resolution, prob = prob, group = group)
  }, 
  depth = dates$depth_corr, 
  prob = likelihoods,
  start = start,
  resolution = resolution,
  group = 1:nrow(dates),
  SIMPLIFY = FALSE)


oxcal <- plyr::ldply(oxcal, I)
oxcal$date <- 1950 - oxcal$date# convert to BP

ggplot(oxcal, aes(x = depth, y = date, colour = prob, group = group, size = prob)) + 
  geom_line() + 
  scale_colour_continuous(low = "#56B1F7", high = "#132B43") + 
  scale_x_reverse() +
  scale_size(range = c(0, 6)) +
  coord_flip()


```

P sequence model building

```{r make_Psequence}
ageDepths <- paste0("R_date(\"Dep", dates$depth_corr, "\",", dates$raw_14C_age, ",", dates$error, "){z=", dates$depth_corr, "};")
ageDepths <- rev(ageDepths)#oldest first
ageDepths <- paste(ageDepths, collapse = "\n")
commandp <- paste(curve, DR, "P_Sequence(\"k\",1,1,U(-2,2)){\nBoundary(\"Bottom\");", ageDepths,    "Boundary(\"Top\");
  };", sep = "\n")
writeLines(commandp)

writeLines(commandp, con = "oxcalp.input")
system(paste("/home/gbsrt/oxcal/OxCal/bin/OxCalLinux", "oxcalp.input"))

```

Read P sequence model

```{r readPseqmod}

js <- readLines("oxcalp.js")

depths <- get.values(js, "data\\.z")
posterior <- get.vectors(js, "posterior\\.prob=")
comment <- get.values(js, "posterior\\.comment\\[0\\]", FALSE)
comment <- comment[!grepl("MCMC", comment)]

p_start <- get.values(js, "posterior\\.start")
p_resolution <- get.values(js, "posterior\\.resolution")

fortify_oxcal <- function(depths, prob, start, resolution, comment = NA){
   out <- mapply(
    function(depth, prob, start, resolution, group, comment){
      data.frame(depth = depth, date = start + 0:(length(prob) - 1) * resolution, prob = prob, group = group, comment = comment)
    }, 
    depth = depths, 
    prob = prob,
    start = start,
    resolution = resolution,
    group = 1:length(depths),
    comment = comment,
    SIMPLIFY = FALSE
  )
  out <- plyr::ldply(out, I) 
  out$date <- 1950 - out$date
  out
}

pseq <- fortify_oxcal(depths = depths, prob = posterior[-1], start = p_start[-1], resolution = p_resolution[-1], comment = comment[-(1:2)])

head(pseq)

pseq %>%# filter(grepl("Bottom", comment)) %>%
  ggplot(aes(x = depth, y = date, alpha = prob)) + 
  geom_line() + 
#  scale_colour_continuous(low = "#56B1F7", high = "#132B43") + 
  scale_x_reverse() +
  coord_flip()

```

```{r importOxcal}
js <- readLines("oxcalp.js")

validOCD <- js[grep("data\\.z", js)]
validOCD <- gsub("^(ocd\\[\\d+\\])\\..*","\\1", validOCD)

depths <- get.values(js, "\\.data\\.z", "depth")
comment <- get.values(js, "posterior\\.comment\\[0\\]", "comment", FALSE)
depth_comment <- left_join(depths, comment)


p_start <- get.values(js, "posterior\\.start", "start")
p_resolution <- get.values(js, "posterior\\.resolution", "resolution")
pmeta <- full_join(p_start, p_resolution)
pmeta <- left_join(depth_comment, pmeta)
posterior <- get.vectors(js, "posterior\\.prob=")

l_start <- get.values(js, "likelihood\\.start", "start")
l_resolution <- get.values(js, "likelihood\\.resolution", "resolution")
lmeta <- full_join(l_start, l_resolution)
lmeta <- inner_join(depth_comment, lmeta)
likelihood <- get.vectors(js, "likelihood\\.prob=")

posterior2 <- plyr::adply(pmeta, 1, function(r){
  prob <- posterior[[r$lab]]
  date <- r$start + 0:(length(prob) - 1) * r$resolution
  data_frame(lab = r$lab, depth = r$depth, date = date, prob = prob, comment  = r$comment, posterior = TRUE)
})

likelihood2 <- plyr::adply(lmeta, 1, function(r){
  prob <- likelihood[[r$lab]]
  date <- r$start + 0:(length(prob) - 1) * r$resolution
  data_frame(lab = r$lab, depth = r$depth, date = date, prob = prob, comment  = r$comment, posterior = FALSE)
})

allprob <- rbind(posterior2, likelihood2)
allprob$date <- 1950 - allprob$date


allprob %>% filter(grepl("Dep", comment)) %>% 
  ggplot(aes(x = depth, y = date, size = prob, colour = posterior)) + 
           geom_line(alpha = 0.4) + 
           scale_x_reverse() +
           scale_size(range = c(0, 6)) +
           coord_flip() +
           labs(x = "Depth cm", y = "Date yr BP")

allprob %>% filter(posterior == TRUE) %>% 
  ggplot(aes(x = depth, y = date, alpha = prob)) + 
           geom_line() + 
           scale_x_reverse() +
           scale_alpha(range = c(0, 1)) +
           coord_flip() +
           labs(x = "Depth cm", y = "Date yr BP")



